Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-16 write-ups/PLDI/paper.tex
Space.

To enable multitasking through cooperative kernels, the runtime (in
our case, the mega-kernel) must track the state of all
workgroups, i.e.\ whether a workgroup is waiting or computing a kernel,
maintain consistent context states for each kernel, e.g.\ tracking the number
of active workgroups, and provide a
safe way for these states to be modified in response to $\offerfork$/$\offerkill$. We discuss these issues,
and describe the implementation of an efficient resizing barrier.

@@

Enabling rendering of smooth graphics in parallel with execution of irregular algorithms is an important use case for our approach.  Because our prototype implementation is based on a mega-kernel that takes over the entire GPU (see \mysec\ref{sec:implementation}), we cannot assess this use case directly.

As a proxy for genuine graphics rendering, we devised the following
procedure to determine non-cooperative OpenCL workloads that simulate
the computational intensity of various graphics rendering workloads.
%
We designed a synthetic kernel that occupies all workgrouops of a
given GPU for a parameterised length of time $t$.  We wrote a host
program to repeatedly launch this kernel in an infinite loop.  We then
searched for a maximum value for the parameter $t$ that allowed the
synthetic kernel to execute without having an observable impact on
graphics rendering, where ``observable'' was decided manually by one
of the paper authors.  Having found the appropriate value for $t$, we
ran the host application for $X$ seconds, measuring the time $Y < X$
dedicated to GPU execution during this period, as well as the number
of kernel launches $n$ that were issued.  We used $X \geq 10$ in all
experiments.  The values $(X-Y)/n$ and $X/n$ estimate the average time
spent using the GPU to render the display between kernel calls and the
period at which the OS requires the GPU for display rendering,
respectively.

>>>>

To enable multitasking through cooperative kernels, the runtime (in
our case, the megakernel) must track the state of all
workgroups, i.e.\ whether a workgroup is waiting or computing a kernel,
maintain consistent context states for each kernel, e.g.\ tracking the number
of active workgroups, and provide a
safe way for these states to be modified in response to $\offerfork$/$\offerkill$. We discuss these issues,
and describe the implementation of an efficient resizing barrier.

@@

Enabling rendering of smooth graphics in parallel with execution of irregular algorithms is an important use case for our approach.  Because our prototype implementation is based on a megakernel that takes over the entire GPU (see \mysec\ref{sec:implementation}), we cannot assess this use case directly.

We devised the following
procedure to determine non-cooperative OpenCL workloads that simulate
the computational intensity of various graphics rendering workloads.
%
We designed a synthetic kernel that occupies all workgrouops of a
GPU for a parameterised time period $t$, invoked in an infinite loop by a host application.  We then
searched for a maximum value for $t$ that allowed the
synthetic kernel to execute without having an observable impact on
graphics rendering.  Using the computed value, we
ran the host application for $X$ seconds, measuring the time $Y < X$
dedicated to GPU execution during this period and the number
of kernel launches $n$ that were issued.  We used $X \geq 10$ in all
experiments.  The values $(X-Y)/n$ and $X/n$ estimate the average time
spent using the GPU to render the display between kernel calls and the
period at which the OS requires the GPU for display rendering,
respectively.
